---
title: "Jeffrey_Hw3"
author: "Savannah Jeffrey"
date: "2023-06-18"
output: html_document
---

Question 1
The YAML of a markdown file is responsible for the documents information and formatting, as well as how the file will be knit. 

Question 2
The entire code and output will not show up when the file is knit.

Question 3

3A - Load packages and read in csv

```{r}
library(magrittr)
library(dplyr)
read.csv("bees.csv")
str("bees.csv")
```

3B - Remove year 2021

```{r}
bees <- read.csv("bees.csv")
bees2 <- dplyr::filter(.data = bees, year != "2021")
```

3C - Remove kleptoparasite_bees

```{r}
bees3<- dplyr::select (.data = bees2, - "kleptoparasite_bees")
```


3D - Add column 

```{r}
bees4 <- dplyr::mutate(bees3, total_bee_abundance = rowSums(select(bees3, -year)))
```

3E - Pipe functions  

```{r}
bees %>%
  dplyr::filter(year != "2021") %>%
  dplyr::select ( -"kleptoparasite_bees") %>%
  dplyr::mutate(total_bee_abundance = rowSums(bees[c(1:5,7:8), 2:4])) %>%
  dplyr::ungroup()

```

4A - Load packages and read csv 

```{r}
library(tidyr)
tree_data <- read.csv("tree_lichen.csv")
str(tree_data)
```
4B - Pivot data longer 

```{r}
trees2 <- tree_data %>%
   tidyr::pivot_longer(!"tree", names_to = "lichen_type", values_to = "percent_cover" )
```

4C - Check pivoted data structure 

```{r}
str(trees2)
```


